---
title: "project_B"
author: "Ling"
date: "3/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreign) # for importing spss data
library(GGally) # correlation matrix
library(knitr) # kable
library(tidyverse)
library(openxlsx) # write .xlsx
library(car) # For Anova(), to control the type of sums of squares 

# import data
project_b_dataset <- read.spss("/Users/ling/Downloads/2018_02_25_SPSS_dataset_project_B.sav", to.data.frame=TRUE)

```

```{r export spss files, include = FALSE}
# select relevent variables for mixed-anova in spss (wide format)
RT_ACC_omissions <- project_b_dataset %>%
  select(GEN_nonsmoker, 
         SSRT_neutral, SSRT_general_reward, SSRT_specific_reward, 
         perc_correct_neutral, perc_correct_general_reward, perc_correct_specific_reward,
         perc_omissions_neutral, perc_omissions_general_reward, perc_omissions_specific_reward
  )
  
# output two formats that can be read into spss (didnt work)
write.foreign(as.data.frame(RT_ACC_omissions), "/Users/ling/Downloads/RT_ACC_omissions.txt", "/Users/ling/Downloads/RT_ACC_omissions.sps", package="SPSS")

# output to excel format (work)
write.xlsx(RT_ACC_omissions, "/Users/ling/Downloads/RT_ACC_omissions.xlsx")
```


```{r clean data, include=FALSE}
## tidy datasets for anova in r
# reaction time
RT <- project_b_dataset %>%
  select(RSSFilename, GEN_nonsmoker, SSRT_neutral, SSRT_general_reward, SSRT_specific_reward) %>% 
  gather(key = reward_conditions, value = reaction_time, c(SSRT_neutral,SSRT_general_reward,SSRT_specific_reward), factor_key = TRUE) %>% 
  separate(reward_conditions, into = c("SSRT", "conditions", "reward")) %>% 
  select(RSSFilename, GEN_nonsmoker, conditions, reaction_time)

# accuracy percentage
ACC <- project_b_dataset %>%
  select(RSSFilename, GEN_nonsmoker, perc_correct_neutral, perc_correct_general_reward, perc_correct_specific_reward) %>% 
  gather(key = reward_conditions, value = accuracy, c(perc_correct_specific_reward, perc_correct_neutral, perc_correct_general_reward), factor_key = TRUE) %>% 
  separate(reward_conditions, into = c("perc", "correct", "conditions", "reward")) %>% 
  select(RSSFilename, GEN_nonsmoker, conditions, accuracy)

# ommisions
omissions <- project_b_dataset %>% 
  select(RSSFilename, GEN_nonsmoker,  perc_omissions_neutral, perc_omissions_general_reward, perc_omissions_specific_reward) %>% 
  gather(key = reward_conditions, value = omission, c(perc_omissions_neutral, perc_omissions_general_reward, perc_omissions_specific_reward)) %>% 
  separate(reward_conditions, into = c("perc", "omissions", "conditions", "reward")) %>% 
  select(RSSFilename, GEN_nonsmoker, conditions, omission)

# merge datasets
RT_ACC <- merge(RT, ACC, by = c("RSSFilename", "GEN_nonsmoker", "conditions"))
total_clean <- merge(RT_ACC, omissions, by = c("RSSFilename", "GEN_nonsmoker", "conditions"))

```

## Correlation Matrix

```{r EDA, echo=FALSE}
## EDA
# correlation matrix plots
project_b_dataset %>% 
  select(BIS_Total, STIMP_MEAN, STIMP2_MEAN, POMS_total_mood_disturbance) %>%
  ggpairs()

```

## Descriptive Statistics

```{r descriptive, echo=FALSE}
## descriptive table
descriptive <- total_clean %>%
  group_by(GEN_nonsmoker, conditions) %>% 
  summarise(N = length(RSSFilename),
            Mean_RT = mean(reaction_time, na.rm=TRUE), 
            Mean_ACC = mean(accuracy, na.rm = T),
            Mean_omission = mean(omission, na.rm = T),
            SEM_RT = sd(reaction_time)/sqrt(length(RSSFilename)),
            SEM_ACC = sd(accuracy)/sqrt(length(RSSFilename)),
            SEM_omission = sd(omission)/sqrt(length(RSSFilename)))

kable(descriptive, digits=2, format="pandoc", caption="Descriptive Statistics for reaction time, accuracy and omission")

```

## Plots 

```{r bar plots for 3 DV, echo=FALSE}
# reaction time
descriptive %>% 
  ggplot() +
  aes(x = conditions, y = Mean_RT, fill = as.factor(GEN_nonsmoker)) + 
  geom_bar(position = position_dodge(), stat = "identity", width = .5) +
  geom_errorbar(aes(ymin = Mean_RT - SEM_RT, ymax = Mean_RT + SEM_RT), width = .1, position = position_dodge(0.5)) +
  theme_bw() + 
  scale_fill_discrete(name = "Groups", breaks = c("0", "1"), labels = c("Smokers", "Non-smokers")) +
  ylab("Reaction Time (ms)") +
  xlab("Reward Conditions") +
  ggtitle("Reaction time for different reward conditions\n(bars indicating SEM, N = 26)")
  
# accuracy 
descriptive %>% 
  ggplot() +
  aes(x = conditions, y = Mean_ACC, fill = as.factor(GEN_nonsmoker)) + 
  geom_bar(position = position_dodge(), stat = "identity", width = .5) +
  geom_errorbar(aes(ymin = Mean_ACC - SEM_ACC, ymax = Mean_ACC + SEM_ACC), width = .1, position = position_dodge(0.5)) +
  theme_bw() + 
  scale_fill_discrete(name = "Groups", breaks = c("0", "1"), labels = c("Smokers", "Non-smokers")) +
  ylab("Reaction Time (ms)") +
  xlab("Reward Conditions") +
  ggtitle("Reaction time for different reward conditions\n(bars indicating SEM, N = 26)")

# omission
descriptive %>% 
  ggplot() +
  aes(x = conditions, y = Mean_omission, fill = as.factor(GEN_nonsmoker)) + 
  geom_bar(position = position_dodge(), stat = "identity", width = .5) +
  geom_errorbar(aes(ymin = Mean_omission - SEM_omission, ymax = Mean_omission + SEM_omission), width = .1, position = position_dodge(0.5)) +
  theme_bw() + 
  scale_fill_discrete(name = "Groups", breaks = c("0", "1"), labels = c("Smokers", "Non-smokers")) +
  ylab("Reaction Time (ms)") +
  xlab("Reward Conditions") +
  ggtitle("Reaction time for different reward conditions\n(bars indicating SEM, N = 26)")

```

## Mixed ANOVA  
  **Independent variables**: group (as between); reward conditions (as within).  
  **Dependent variables**: reaction time, accuracy, omission.  
    
### Reactione Time  
   A 2x3 ANOVA to test if there are differences in reaction time based on different group and the types of reward.

```{r mixed anova rt, echo=FALSE}
  # (For ANOVAs with within-subjects variables, the data must be in long format)

# reaction time
# aov_rt <- aov(reaction_time ~ GEN_nonsmoker * conditions + Error(RSSFilename/conditions), data = total_clean)
# summary(aov_rt)

model_rt <- lm(reaction_time ~ GEN_nonsmoker * conditions, data = total_clean)
rt_aov <- Anova(model_rt, type = 3)

# Effect size (partial eta-squared)
rt_aov$pes <- c(rt_aov$'Sum Sq'[-nrow(rt_aov)], NA)/(rt_aov$'Sum Sq' + rt_aov$'Sum Sq'[nrow(rt_aov)]) # SS for each effect divided by the last SS (SS_residual)

# output result
  # change column names and row names
colnames(rt_aov) <- c("ss", "df", "$F$", "$p$", "partial $\\eta^2$") 
rownames(rt_aov) <- c("(Intercept)", "group", "reward conditions", "group:reward conditions", "Residuals") 
  # output table, omit the intercept row
kable(rt_aov[-1, ], digit = 2, format = "pandoc", caption = "ANOVA of Reaction Time")


# table of means
# model.tables(aov_rt, "means")
  # table the results
# kable(aov_rt, digits = 2, format="pandoc", caption="ANOVA of Reaction Time")
# 
# library(apaTables)
# apa.aov.table(aov_rt, filename = "RT table.doc")

```

### Accuracy  

   A 2x3 ANOVA to test if there are differences in accuracy based on different group and the types of reward.

```{r mixed anova acc, echo=FALSE}
# fit a model, accuracy
model_acc <- lm(accuracy ~ GEN_nonsmoker * conditions, data = total_clean)
acc_aov <- Anova(model_acc, type = 3)
 # effect size
acc_aov$pes <- c(acc_aov$'Sum Sq'[-nrow(acc_aov)], NA)/(acc_aov$'Sum Sq' + acc_aov$'Sum Sq'[nrow(acc_aov)])
 # output
colnames(acc_aov) <- c("ss", "df", "$F$", "$p$", "partial $\\eta^2$") 
rownames(acc_aov) <- c("(Intercept)", "group", "reward conditions", "group:reward conditions", "Residuals") 
kable(acc_aov[-1, ], digit = 2, format = "pandoc", caption = "ANOVA of Accuracy")

```


### Omission  

   A 2x3 ANOVA to test if there are differences in omission based on different group and the types of reward.

```{r mixed anova omission, echo=FALSE}
# fit a model, omission
model_omi <- lm(omission ~ GEN_nonsmoker * conditions, data = total_clean)
omi_aov <- Anova(model_omi, type = 3)
 # effect size
omi_aov$pes <- c(omi_aov$'Sum Sq'[-nrow(omi_aov)], NA)/(omi_aov$'Sum Sq' + omi_aov$'Sum Sq'[nrow(omi_aov)])
 # output
colnames(omi_aov) <- c("ss", "df", "$F$", "$p$", "partial $\\eta^2$") 
rownames(omi_aov) <- c("(Intercept)", "group", "reward conditions", "group:reward conditions", "Residuals") 
kable(omi_aov[-1, ], digit = 2, format = "pandoc", caption = "ANOVA of Omission")
```




